---
title: "Network sensitivity analysis for TB-assortativity"
output: pdf_document
editor_options: 
  chunk_output_type: console
header-includes: 
- \usepackage[section]{placeins}
- \usepackage{amsmath}
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(readr)
library(magrittr)
library(tidyr)
library(igraph)
library(ggplot2)
library(cowplot)
library(plotrix)
library(knitr)
library(colorspace)
library(tidyverse)
library(gridExtra)
library(ggraph)

knitr::opts_chunk$set(cache=TRUE, fig.path = "slirs-par-figs/", 
                      warning = FALSE, message=FALSE, echo=FALSE, fig.height =6, fig.width = 8) 

# colors for assortativity
assortCols <- c("#F3E79AFF", "#F7A086FF", "#CF63A6FF", "#704D9EFF")

```


# How does the way we generate assorted networks affect results? 

```{r}

res <- read.csv("simulations-sah/res_sah_50.csv")

res$modNameLatent <- ifelse(res$reactivation_rate==0.1, T, F)
res$modNameRecover <- ifelse(res$reversion_rate==0.33, T, F)
res$modName <- NA
res$modName[!res$modNameLatent&!res$modNameRecover] <- "SIR"
res$modName[res$modNameLatent&!res$modNameRecover] <- "SLIR"
res$modName[!res$modNameLatent&res$modNameRecover] <- "SIRS"
res$modName[res$modNameLatent&res$modNameRecover] <- "SLIRS"

res$alph_type <- factor(res$alph_type, labels = c("IP", "SUS", "TRA"))

res$mf_r_ratio <- res$m_rec/res$f_rec
res$mf_i_ratio <- res$m_inf/res$f_inf

res$r <- res$r * 2 # sah uses modularity, so assortaaivity is multipled by 2

```



```{r prev-rat-barplot, fig.cap="The effects of sex-traits and assortativity on male-bias in simulations on Sah-networks. Faceted by transmission rate. Results shown for combined sex-traits (IP, SUS, TRA) ", fig.height=8}

res %>%
  filter(modName %in% c("SIR", "SLIR")) %>%
  select(modName, r, tau, alph_type, alph_val, mf_r_ratio, duration, 
         net_clustering, net_path_len, net_deg_assort) %>%
  rename(mf_ratio=mf_r_ratio)-> mods1

res %>%
  filter(modName %in% c("SIRS", "SLIRS")) %>%
  select(modName, r, tau, alph_type, alph_val, mf_i_ratio, duration, 
         net_clustering, net_path_len, net_deg_assort) %>%
  rename(mf_ratio=mf_i_ratio)-> mods2

comRes <- rbind(mods1, mods2)

comRes %>%
  ggplot(aes(x=factor(r), y=mf_ratio, fill=factor(alph_val))) +
  geom_boxplot()  +  facet_grid(tau~modName) +
  scale_y_log10() +
  labs(x="Assortativity (r)", y=("Male:Female Infection Ratio"), 
       fill="Alpha", shape="Model \n type") +
  theme_bw() + ggtitle("IP, SUS, TRA") + 
  theme(axis.text=element_text(size=14),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.title=element_text(size=14,face="bold"),
        legend.title = element_text(size=14),
        legend.text = element_text(size=14), 
        # panel.grid.major = element_blank(),
        # panel.grid.minor = element_blank(), 
        strip.text = element_text(size = 14, color="Black")) 

```



```{r prev-rat-heatmap-low-tau, fig.cap="The effects of sex-traits and assortativity on male-bias in simulations on Sah-networks. Results shown for low transmissability pathogen. Only ratios above 1.25 are shown to visualize which parameter combinations can lead to male-bias. ", fig.height=8}

hs <- 14 # text size
 
res %>%
  filter(!is.infinite(mf_r_ratio), tau==0.075, modName%in%c("SLIR", "SIR")) %>%
  group_by(modName, net_type, alph_type, r, alph_val) %>%
  summarise(meanPrev=mean(mf_r_ratio, na.rm=TRUE)) -> fooa

res %>%
  filter(!is.infinite(mf_i_ratio), tau==0.075, modName%in%c("SLIRS", "SIRS")) %>%
  group_by(modName, net_type, alph_type, r, alph_val) %>%
  summarise(meanPrev=mean(mf_i_ratio, na.rm=TRUE)) -> foob

# for SLIR use R, for SLIRS use I class for MF calculation
foo <- bind_rows(fooa, foob)

foo %>%
  mutate(absPrev=abs(meanPrev-1.8)) %>%
  group_by(net_type, modName, alph_type, r) %>%
  summarise(alphVal=which.min(absPrev)) -> foo2

foo %>%
  ggplot(aes(x=factor(r), y=factor(alph_val))) +
  geom_tile(aes(fill=meanPrev)) + 
  scale_fill_gradient2(midpoint=log10(1.8), trans="log10", 
                                     low="red4", mid = "dodgerblue",
                       high="black", limits=c(1.25,max(foo$meanPrev)), na.value = "white") +
  # geom_point(data=foo2, aes(x=factor(r), y=alphVal), 
  #            shape=17, color="white") +
  facet_grid(modName ~ alph_type) +
  labs(x="Assortativity (r)", y=expression(alpha), 
       fill="Mean \nPrevalence \nRatio", title = "R0=1.5") +
  theme_bw() +
  theme(axis.text=element_text(size=hs),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.title=element_text(size=hs,face="bold"),
        legend.title = element_text(size=hs),
        legend.text = element_text(size=hs), 
        # panel.grid.major = element_blank(),
        # panel.grid.minor = element_blank(), 
        strip.text = element_text(size = 18, color="Black")) 

```

\clearpage

```{r outbreak-compare1, fig.cap="Epidemic size in SLIR model simulations on Sah networks across assortativity values. Faceted by transmission rate. Colors show sex-trait ratios. "}

foo2 <- res %>% 
  filter(modName%in% c("SIR", "SLIR"), alph_val==1) %>%
  mutate(out_size=(m_rec + f_rec)/1000) %>%
  group_by(modName, tau, r) %>%
  summarise(out_size=mean(out_size))


res %>%
  filter(modName=="SLIR", tau %in% c(0.04, 0.15)) %>%
  mutate(size=(f_rec + m_rec)) %>%
  ggplot(aes(x=factor(r), y=size, fill=factor(alph_val))) +
  geom_boxplot()  +  facet_wrap(tau~.) +
  #scale_color_continuous_diverging() +
  labs(x="Assortativity (r)", y=("Epidemic size"), 
       fill="Alpha", shape="Model \n type") +
  theme_bw() + ggtitle("SLIR") + 
  theme(axis.text=element_text(size=14),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.title=element_text(size=14,face="bold"),
        legend.title = element_text(size=14),
        legend.text = element_text(size=14), 
        # panel.grid.major = element_blank(),
        # panel.grid.minor = element_blank(), 
        strip.text = element_text(size = 14, color="Black")) 

```


```{r duration-compare1, fig.cap="Epidemic duration in SLIR model simulations on Sah networks across assortativity values. Faceted by transmission rate. Colors show sex-trait ratios. "}

res %>%
  filter(modName=="SLIR", tau %in% c(0.04, 0.15)) %>%
  ggplot(aes(x=factor(r), y=duration, fill=factor(alph_val))) +
  geom_boxplot()  +  facet_wrap(tau~.) +
  #scale_color_continuous_diverging() +
  labs(x="Assortativity (r)", y=("Epidemic duration"), 
       fill="Alpha", shape="Model \n type") +
  theme_bw() + ggtitle("SLIR") + 
  theme(axis.text=element_text(size=14),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.title=element_text(size=14,face="bold"),
        legend.title = element_text(size=14),
        legend.text = element_text(size=14), 
        # panel.grid.major = element_blank(),
        # panel.grid.minor = element_blank(), 
        strip.text = element_text(size = 14, color="Black")) 

```


```{r outbreak-compare2, fig.cap="Endemic prevalence in SLIRS model simulations on Sah networks across assortativity values. Faceted by transmission rate. Colors show sex-trait ratios. ", fig.width=8, fig.height=7}


res %>%
  filter(modName=="SLIRS", tau %in% c(0.04, 0.15)) %>%
  mutate(infected_prev=(m_inf+f_inf)/1000) %>%
  ggplot(aes(x=factor(r), y=infected_prev, fill=factor(alph_val))) +
  geom_boxplot()  +  facet_wrap(tau~.) + ggtitle("SLIRS") +
  #scale_color_continuous_diverging() +
  labs(x="Assortativity (r)", y=("Endemic prevalence"), 
       fill="Alpha", shape="Model \n type") +
  theme_bw() +
  theme(axis.text=element_text(size=14),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.title=element_text(size=14,face="bold"),
        legend.title = element_text(size=14),
        legend.text = element_text(size=14), 
        # panel.grid.major = element_blank(),
        # panel.grid.minor = element_blank(), 
        strip.text = element_text(size = 14, color="Black")) 

```

\clearpage

## Sensitivity analyses 

```{r sah-structure, fig.cap="The effects of assortativity on network structural characteristics in Sah networks. "}

res %>%
  gather("stat", "val", c(net_clustering, net_path_len, net_deg_assort)) %>%
  ggplot(aes(factor(r), val)) + geom_boxplot() + facet_grid(stat~., scales="free")
  


```


```{r sah-structure-bias, fig.cap="The relationship between network structural characteristics and male-bias. Results shown for all model types (SIR, SLIR, SIRS, SLIRS) and sex-traits. "}

comRes %>% 
  #filter(alph_type=="SUS")%>%
  gather("stat", "val", c(net_clustering, net_path_len, net_deg_assort)) %>%
  ggplot(aes(val, mf_ratio, color=factor(r))) + 
  #geom_point(size=.05) +
  stat_ellipse() +
  facet_grid(.~stat, scales="free_x") +
  theme_bw() +
  theme(axis.text=element_text(size=14),
        axis.text.x = element_text(hjust=1, angle=45),
        axis.title=element_text(size=14,face="bold"),
        legend.title = element_text(size=14),
        legend.text = element_text(size=14), 
        # panel.grid.major = element_blank(),
        # panel.grid.minor = element_blank(), 
        strip.text = element_text(size = 14, color="Black")) 

```

* Preliminary results for 50 replicates on networks generated with Sah algorithm (r=0, .3, .6, .9)
* One epidemic simulated per network
* Simulations stopped after 150 time steps, need to increase reps & length of simulation

* Results largely similar to re-wiring. Assortativity doesn't inrease male-bias alone. Sex-traits, especially longer male infecitous period, are likely the culprit. Any differences found in outbreak size are likely due to changes in network statistics, including changing clustering, degree assortativity, and network path length rather than assortativity. 



