---
title: "Making assorted networks by re-wiring"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(magrittr)
library(tidyr)
library(igraph)
library(ggplot2)
library(cowplot)
library(plotrix)
library(knitr)
library(igraph)

knitr::opts_chunk$set(cache=FALSE, fig.path = "assort-nets-op2/", 
                      warning = FALSE, message=FALSE, echo=FALSE) 

```

We will generate scale-free networks according to the parameters listed in Table 1 using the classic BA-algorithm. Following network generation, we will update the networks as following: 

1. Assign nodes randomly as male (0) or female (1). 
2. Calculate temporary value of sex-assortativity in the network ($r_t$). 
3. If $r_t$ is not within $\epsilon$ of $r$, randomly choose a proportion $\alpha$ of 0--1 edges (i.e, a male--female edge) if $r\geq 0$ and re-wire them or if $r < 0$, choose a proportion $\alpha$ of 0--0 and 1--1 edges and re-wire the. 
4. Repeat step 3 until $|r_f - r_t|\leq \epsilon$. 

| Variable  | Value  | 
|:-:|:-:|
| Sex-assortativity, $r$  |  -0.3, -0.2, - 0.1, 0, 0.1, 0.2, 0.3 |
| Degree distribution, $p(k)$ | $\frac{k^{-\alpha}}{\zeta (\alpha)}$  |
| Mean degree, $<k>$ | 10  |
| Network size, $N$ | 500, $1\cdot 10 ^ 3$, $1.5\cdot 10 ^ 3$  |
| Tolerance, $\epsilon$ | 0.035  |
| Rewiring proportion, $\alpha$ | 0.2  |

Table: Design of pilot study to test rewiring algorithm.   

```{r}
setwd("~/Documents/phd/research-projects/miller-tb-assortativity/analysis/network-generation/op-2-simulations/")

algDat <- data.frame(size=NA, r=NA, rep=NA, degAssort=NA, clustering=NA, pathLen=NA)

parms=expand.grid(size=c(5e2, 1e3, 1.5e3), 
                  r=c(-0.3, -0.2, -0.1, 0, 0.1, 0.2, 0.3), 
                  rep=1:15)

for(i in 1:nrow(parms)){
  s=parms[i, "size"]
  r=parms[i, "r"]
  rep=parms[i, "rep"]
  
  net=paste0("G_", as.character(r), "N", s, "rep", rep)
  g=read.graph(paste0(net, ".graphml"), format = "graphml")
  
  foo <- c(s, r, rep, assortativity_degree(g, directed = FALSE),
           transitivity(g), diameter(g, directed = FALSE))
  algDat[i, ] <- foo
}

```

```{r algCompare, fig.cap="Values of clustering, degree assortativity, and path length with varying assortativity for networks generated using rewiring algorithm. "}

algDat %>%
  select(-rep) %>%
  gather(stat, value, 3:5) %>%
  ggplot(aes(x=r, y=value, color=factor(size))) + 
  geom_smooth() +
  geom_point() + 
  facet_grid(stat ~ ., scales="free") + 
  ylab("") + xlab("Assortativity (r)")
  
```


```{r exampleNets, fig.height=7, fig.cap="Networks generated using rewiring algorithm (N=500) with corresponding edge and degree distributions. Edge distribution shows number of edges within module (sex) and between module (sex). "}
plot_assort_graph <- function(graph){
  
  ##basic network layout
  nn <- 1
  types <- V(graph)$sex
  xy1 <- c(-1.5, .1) # males
  xy2 <- c(1.5, -.1) # females
  
  get_coords <- function(tps, loc1, loc2){
    # function that takes in type and loc of types and 
    # returns assignedn location with noise based on type
    
    if(tps==1){
      x=rnorm(1, mean=loc1[1], sd=.65)
      y=rnorm(1, mean=loc1[2], sd=.65)
    }
    if(tps==2){
      x=rnorm(1, mean=loc2[1], sd=.65)
      y=rnorm(1, mean=loc2[2], sd=.65)
    }
    return(c(x,y))
  }
  
  coords=lapply(types, get_coords, loc1=xy1, loc2=xy2)
  coords=do.call(rbind, coords)
  
  plot(graph, layout=coords, 
       vertex.label="", 
       vertex.size=3, vertex.color=V(graph)$sex, 
       edge.width=.1, edge.color="black")
}
get_e_type <- function(elRow, types){
  ifelse(diff(types[elRow])==0, "w", "b")
}

setwd("~/Documents/phd/research-projects/miller-tb-assortativity/analysis/network-generation/op-2-simulations/")

# plots of network generation process
layout(matrix(1:9, 3, 3, byrow = TRUE))
#layout.show(n=9)

# DISASSORTATIVE
par(mar=c(1,1,2,1))
net=paste0("G_", as.character(-0.3), "N", 500, "rep", 1)
gD=read.graph(paste0(net, ".graphml"), format = "graphml")
plot_assort_graph(gD)
rt0 <- assortativity_nominal(gD, types=V(gD)$sex) # basically 0
mtext(round(rt0, 3))
el <- get.edgelist(gD)
eTypes <- apply(el, 1, get_e_type, types=V(gD)$sex)
par(mar=c(3,3,3,3))
barplot(table(eTypes))
par(mar=c(3,3,3,3))
plot(degree_distribution(gD))

# Random
par(mar=c(1,1,2,1))
net=paste0("G_", as.character(0), "N", 500, "rep", 1)
gR=read.graph(paste0(net, ".graphml"), format = "graphml")
plot_assort_graph(gR)
rt0 <- assortativity_nominal(gR, types=V(gR)$sex) # basically 0
mtext(round(rt0, 3))
el <- get.edgelist(gR)
eTypes <- apply(el, 1, get_e_type, types=V(gR)$sex)
par(mar=c(3,3,3,3))
barplot(table(eTypes))
par(mar=c(3,3,3,3))
plot(degree_distribution(gR))

# ASSORTATIVE
par(mar=c(1,1,2,1))
net=paste0("G_", as.character(0.3), "N", 500, "rep", 1)
gA=read.graph(paste0(net, ".graphml"), format = "graphml")
plot_assort_graph(gA)
rt0 <- assortativity_nominal(gA, types=V(gA)$sex) # basically 0
mtext(round(rt0, 3))
el <- get.edgelist(gA)
eTypes <- apply(el, 1, get_e_type, types=V(gA)$sex)
par(mar=c(3,3,3,3))
barplot(table(eTypes))
par(mar=c(3,3,3,3))
plot(degree_distribution(gA))

```

Notes: 

* 15 replicates of each size and assortativity
* Takes less than one hour to run
* Algorithm needs to be re-run to remove multiple edges

